---
output: 
  github_document:
    toc: TRUE
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "50%"
)
```


```{r, echo = F, eval = F}
library(ggplot2)
library(ggstamp)

ggbarlabs() +
  stamp_polygon_inverse(radius_outer = Inf) +
  stamp_polygon(color = "darkred", 
                fill = "lightgoldenrod1",
                alpha = 0,
                size = 3) + 
  # stamp_arrow(x = -.75, xend = .75,
  #             y = 0, yend = 0,
  #             size = 4, 
  #             headlength = .2, 
  #             angle = 20,
  #             color = "darkseagreen", 
  #             type = "closed"
  #             ) + 
  stamp_text(#y = .35, x = -.25, 
    x = -.3, y = -.34,
             label = "ggbarlabs", 
             size = 12, 
             color = "dodgerblue4",
             vjust = 0) +
  theme_void()
```

# ggbarlabs

<!-- badges: start -->
<!-- badges: end -->

# Problem: 

## bar charts are ubiquitous and can quickly communicate information.

```{r}
library(ggplot2)
theme_set(theme_gray(18))
ggplot(mtcars) + 
  aes(x = am) + 
  geom_bar() 

p <- last_plot()

```


## bar plots can benefit from specificity of labeling

Labeled bar chart is all the fast communication of traditional data vizualization with all the specificity of a data table.

```{r}



```


## but its a pain 

either precalc and use geom_col + geom_text

```{r}


```

## or use verbose after_stat...


### first inspecting bar layer (stat_count)

```{r}
layer_data(last_plot(), 1)
```

### then plot

using our knowledge of what data frame results when using StatCount, refer to the computed var, count

```{r}
# count column can be used via after_stat
p +
  geom_text(stat = StatCount, 
            aes(label = after_stat(count)), 
            vjust = -.7)

p + 
  geom_text(stat = StatCount, 
            aes(label = after_stat(
              paste0(round(100*count/sum(count), 1) ,"%")
              )), 
            vjust = -.7)

p + 
  geom_text(stat = StatCount, 
            aes(label = after_stat(
              paste0(count,"\n(",round(100*count/sum(count), 1) ,"%)")
              )), 
            vjust = -.5, lineheight = .8)

layer_data(last_plot(), 2)
```



## What if... ggbarlabs Proposed User interface


```{r, eval = F}
ggplot(mtcars) + 
  aes(x = am) + 
  geom_bar() + 
  geom_text_count()

ggplot(mtcars) + 
  aes(x = am) + 
  geom_bar() + 
  geom_text_count_percent()
```


## Composing functions

### geom_text_count


```{r geom_text_count}
#' geom_text_count
#' @description Add label which is the count.  Acts like geom_bar() but text label at the position of the top of the bar.
#'
#' @param nudge_y
#' @param position
#' @param ...
#'
#' @return
#' @export
#'
#' @examples
geom_text_count <- function(nudge_y = 0, position =
                              ggplot2::position_dodge2(width = .9,
                                                       preserve = "single"), ...){

  ggplot2::stat_count(geom = "text",
                      ggplot2::aes(label = ggplot2::after_stat(count),
                                   y = ggplot2::after_stat(count) + nudge_y),
                      vjust = 0,
                      position = position,
                      ...
  )

}
```


### Test it out

```{r}
library(ggplot2)
ggplot(mtcars) +
   aes(x = factor(cyl)) +
   geom_bar(position = "dodge") +
   geom_text_count(nudge_y = .2)

last_plot() + 
  aes(fill = factor(am))
```


### geom_text_count_percent

```{r}


#' geom_text_count_percent
#' @description Add label which is the count as well as percentage.  Acts like geom_bar() but text label at the position of the top of the bar.  Percentage is calculated within facet panels.
#'
#' @param nudge_y
#' @param lineheight
#' @param position
#' @param ...
#'
#' @return
#' @export
#'
#' @examples
geom_text_count_percent <- function(nudge_y = 0,
                                    lineheight = .85,
                                    position = ggplot2::position_dodge2(width = .9,
                                                               preserve = "single"), ...){

  stat_count(geom = "text",
             ggplot2::aes(label = paste0(after_stat(count), "\n(",

                                round(
                                  100*(ggplot2::after_stat(count))/
                                    tapply(ggplot2::after_stat(count),
                                           ..PANEL..,
                                           sum)[..PANEL..],
                                  1), "%)" ),
                 y = ggplot2::after_stat(count) + nudge_y),
             vjust = 0,
             lineheight = lineheight,
             position = position,
             ...
             )
}
```


### Try it out

```{r}
library(ggplot2)
ggplot(mtcars) +
    aes(x = factor(cyl)) +
    geom_bar(position = "dodge") +
    geom_text_count_percent(nudge_y = .25)

last_plot() +
    aes(fill = factor(am))
```


# Furthermore, a different set of thematic and scale defaults make sense for labeled bar charts


```{r}
ggplot(mtcars) +
  aes(x = factor(cyl)) +
  geom_bar(position = "dodge") +
  geom_text_count(nudge_y = .2) +
  theme_classic() +
  theme(axis.line.y = element_blank(),
        axis.text.y.right = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        panel.grid.major.y = element_line(color = alpha("gray35", .25)),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        axis.line.x = element_line(colour = "gray35"),
        legend.position = "top",
        legend.justification = 0) +
  scale_y_continuous(expand = expansion(mult = c(0, .1)))
```

# lets put all of this in `defaults_ggbarlabs` and `ggbarlabs()`


```{r}
defaults_ggbarlabs <- function(){
  
  list(
  theme_classic(base_size = 15) ,
  theme(axis.line.y = element_blank(),
        axis.text.y.right = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        panel.grid.major.y = element_line(color = alpha("gray35", .25)),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        axis.line.x = element_line(colour = "gray35"),
        legend.position = "top",
        legend.justification = 0) ,
  scale_y_continuous(expand = expansion(mult = c(0, .15))))
}


ggbarlabs <- function(data = NULL, ...){
  ggplot(data= data , ... ) +
  defaults_ggbarlabs()
}
```

## try it out

```{r}
ggplot(mtcars) + 
  aes(x = factor(am)) + 
  geom_bar(fill = alpha("navy", .9)) + 
  geom_text_count_percent() + 
  defaults_ggbarlabs()

# or
ggbarlabs(mtcars) + 
  aes(x = factor(am), fill = factor(cyl)) + 
  geom_bar(position = "dodge") + 
  geom_text_count_percent()


```




## Issues

### vjust may give us better results than nudge_y... 
### move ggbarlabs()
### percents is calculated within panel.  We might want to specify the 'whole' from which percentage is calculated.
### horizontal bars

